---
title: "Homework #4"
author: "Qi Tan"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: yes
    toc: true
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{homework-4 vignette}
-->

# 1. CASL Number 2 in Exercises 5.8.
Linear Hessian: $H=X^tX$\
Hessian matrix: $H=X^tdiag(p(1-p))X$\
According to formula 2.72, condition number a matrix is the ratio of its largest and smallest singular values, $$cond(A)=\frac{\sigma_{max}}{\sigma_{min}}$$
```{r}
X <- matrix(c(2, 1, 1, 2), 2, 2) #create a matrix for X
beta <- c(0, 10)
p <- 1/(1+exp(-X %*% beta))
p

#Calculate condition number of linear Hessian
H1 <- t(X) %*% X
svals1 <- svd(H1)$d
cond_hs <- max(svals1)/min(svals1)
cond_hs #9

#Calculate condition number of logistic variation
W <- as.numeric(p*(1-p))
H2 <- crossprod(X, diag(W) %*% X)
svals2 <- svd(H2)$d
cond_lg <- max(svals2)/min(svals2)
cond_lg #61182.63
```

Linear Hessian is well-conditioned since it has a low condition number, while logistic variation is ill-conditioned with a high condition number.



# 2. CASL Number 4 in Exercises 5.8. 
Using closed-form solution for ridge regression: $$\hat{\beta}^{(\lambda)}=(X^TX+\lambda I_p)^{-1}X^Ty$$

```{r}
# Solve generalized linear models with Newton-Ralphson method.
#
# Args:
#     X: A numeric data matrix.
#     y: Response vector.
#     family: Instance of an R ‘family‘ object.
#     maxit: Integer maximum number of iterations.
#     tol: Numeric tolerance parameter.
#
# Returns:
#     Regression vector beta of length ncol(X).
casl_glm_irwls <- function (X, y, family, maxit=25, tol=1e-10, lambda=1e-4) {
  beta <- rep(0, ncol(X))
  for (j in seq_len(maxit)) {
    b_old <- beta
    eta <- X %*% beta
    mu <- family$linkinv(eta)
    mu_p <- family$mu.eta(eta)
    z <- eta + (y - mu) / mu_p
    W <- as.numeric(mu_p^2 / family$variance(mu))
    XtX <- crossprod(X, diag(W) %*% X)
    Xtz <- crossprod(X, W * z)
    I <- diag(ncol(XtX)) #Create I_p
    R <- lambda*I #lambda*I_p
    beta <- solve(XtX + R, Xtz)
    if (sqrt(crossprod(beta - b_old)) < tol) break
  }
  beta 
}
```
